---
title: "ESM 244 Assignment 1 Task 1"
author: "Kirsten Hodgson"
date: "1/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)

library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(readxl)
library(lubridate)
```

#### Read-in data and set-up

```{r}
# Read in data and initial wrangling
#Use read_excel to read in data from an excel spreadsheet
sierra_amphibians <- read_excel(here("data","sierra_amphibians.xlsx")) 

#While preserving the original data, create a subset with some wrangling
amphibians_clean <- sierra_amphibians %>% 
  clean_names() %>% #convert variables to lower snake
  mutate(across(where(is.character), tolower)) %>% #character obs lowercase
  mutate(survey_date = ymd(survey_date)) %>% #change the year to a date class
  mutate(year = year(survey_date)) #Create a new column with just year
```


#### Task 1a: Mountain yellow-legged frog counts by life stage and year

```{r}
#Create another subset for R. muscosa, then get counts
r_muscosa_counts <- amphibians_clean %>% 
  filter(amphibian_species == "ramu") %>% #Filter by species
  filter(amphibian_life_stage != "eggmass") %>% #Remove eggmass life stage
  group_by(year) %>% 
  count(amphibian_life_stage, wt = amphibian_number) #Count R. muscosa obs by year, life stage
```

```{r, fig.width = 12}
#Create a character vector to use for labeling facets later
facet_labels <- c(`adult` = "Adult",
                  `subadult` = "Sub-Adult",
                  `tadpole` = "Tadpole")
#Create a plot of counts and customize
counts_by_year <- ggplot(data = r_muscosa_counts, 
       aes(x = year, y = n)) +
  geom_point(size = 2.3, color = "darkgreen") + #Customize size and color of points
  labs(x = "\nYear", y = "Count\n") + #Add axis labels
  facet_wrap(~amphibian_life_stage, #Facet by life stage
             scales = "free_y", #With free y-axis scales to account for very different numbers
             labeller = as_labeller(facet_labels)) + #And the labels from the earlier vector
  theme_minimal() +
  theme(legend.position = "none") #Remove legend
```


#### Task 1b: Lakes with the highest abundance of mountain yellow-legged frog adults and subadults

```{r}
#Wrangling for task 1b
lakes_r_muscosa <- amphibians_clean %>% 
  filter(amphibian_species == "ramu", #Keep only observation of rana muscosa
         amphibian_life_stage %in% c("adult", "subadult")) %>% #Which are also adult or subadult
  mutate(lake_id = paste("Lake", lake_id, sep = " ")) %>% #Add "Lake" before id number
  count(lake_id, wt = amphibian_number) %>% #Count the total number of obs per lake for both stages
  rename(r_muscosa_count = n) %>% #Change name of count column
  slice_max(r_muscosa_count, n = 5) %>% #Keep only the top 5 lakes
  mutate(lake_id = fct_reorder(lake_id, desc(r_muscosa_count))) #Reorder the lake id so in order of descending r muscosa count
```

```{r}
#Create a graph for task 1b
counts_by_lake <- ggplot(data = lakes_r_muscosa, aes(x = lake_id, y = r_muscosa_count)) +
  geom_col(fill = "darkgreen", width = 0.6) + #Column graph with dark green narrower bars
  scale_y_continuous(limits = c(0,2500), #Set the limits in the y direction
                     breaks = c(0,500,1000,1500,2000,2500), #With specified breaks
                     expand = c(0, 0.05)) + #And shorter distance between 0 and x-axis labels
  labs(x = "\nLake", y = "Total Adult and Sub-adult Count\n") +
  theme_minimal()
```


#### Task 1c: Creating a compound figure with `patchwork`

```{r, fig.width = 10, fig.height = 8}
r_muscosa_fig <- ((counts_by_year + theme(plot.margin = unit(c(0,0,20,0), "pt")))/counts_by_lake)
#Arrange the graphs over each other, with some margin added to the top plot to increase spacing
r_muscosa_fig
```

**Figure 1. Counts of mountain yellow-legged frog (*Rana muscosa*).** The top figure shows *R. muscosa* by life stage (adult, sub-adult, and tadpole) and year (1995-2002) in all observed water bodies in the Southern Sierra Nevada. The bottom figure shows total count over all study years of adult and sub-adult *R. muscosa* in the five lakes with the highest *R. muscosa* counts. Data source: Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3
